\Title

### 2013-02-13 supplementary.tex
text small changes

\newpage
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Synonymous diversity across the HIV genome}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h]
\begin{center}
\includegraphics[width=\linewidth]{conservation_codons_genome}
\caption{
Synonymous diversity across the HIV genome, as quantified by the normalized
codon entropy among sequences coding for the consensus amino-acid. The
synonymous diversity peaks at the variable regions in {\it env} and is reduced
in regions under purifying selection (RRE hairpin, second {\it tat}/{\it rev}
exons). Since the variable regions are under positive selection at the protein
level, synonymous hitchhiking is stronger there, and even deleterious synonymous
mutations can be brought to high frequency. This gives to our signal in the
fixation probability, Figs. 2A and 2B of the main text. The normalized codon
entropy is calculated as follows (see the script for the full algorithm): (i)
from a subtype B multiple sequence alignment (MSA) from the LANL website
(filtered sequences only, version 2011)~\cite{LANL2012}, we calculate the
consensus amino acid at each position in the HIV genome; (ii) we count how often
each codon coding for the consensus amino acid appears in the MSA; (iii) at each
amino acid position, we divide by the number of sequences in the MSA that had
the consensus amino acid at that position, obtaining {\it codon frequencies}
$\nu_c$; (iv) we calculate the codon entropy from each position as: $S := -
\sum_{c} \nu_c \log \nu_c$, where $c$ runs over codons that code for the
consensus amino acid at this site; (v) we divide by the maximal codon entropy of
that amino acid (e.g. $\log 2$ for twofold degenerate codons). All parts of
{\it env} that are part of a different gene (signaling peptide, second {\it rev}
exon) have been excluded from our main analysis, to avoid contamination about
synonymity. Note that all gap-rich columns of the MSA are stripped from this
figure, hence genes such as {\it env} might appear shorter than they actually
are.
}
\label{fig:syndiv_genome}
\end{center}
\end{figure}
\newpage
% 
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section{Nonsynonymous changes outside of variable regions are deleterious}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=0.8\linewidth]{synmut_conservation_4fold_synnonsyn}
% \caption{Cumulative distribution of synonymous and nonsynonymous diversity in
% {\it gag} in the LANL reference panel (filtered
% sequences only, version 2011)~\cite{LANL2012}. Sites such that the consensus codon
% has three synonymous and six nonsynonymous single mutants were used, and the number of observed mutants
% of a certain type over the number of possible mutants is plotted. Nonsynonymous
% changes are observed less often, {\it ergo} are more conserved, than
% synonymous changes. It can be therefore assumed, as mentioned in the main text,
% that non-escape nonsynonymous changes involve a large fitness cost.}
% \label{fig:synnonsyncons}
% \end{center}
% \end{figure}
% \newpage

>>>>

\newpage
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Synonymous diversity across the HIV genome}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h]
\begin{center}
\includegraphics[width=\linewidth]{conservation_codons_genome}
\caption{
Synonymous diversity across the HIV genome, as quantified by the normalized
codon entropy among sequences coding for the consensus amino acid. The
synonymous diversity peaks at the variable regions in {\it env} and is reduced
in regions under purifying selection (RRE hairpin, second {\it tat}/{\it rev}
exons). Since the variable regions are under positive selection at the protein
level, synonymous hitchhiking is stronger there, and even deleterious synonymous
mutations can be brought to high frequency. This gives to our signal in the
fixation probability, Figs. 2A and 2B of the main text. The normalized codon
entropy is calculated as follows (see the script for the full algorithm): (i)
from a subtype B multiple sequence alignment (MSA) from the LANL website
(filtered sequences only, version 2011)~\cite{LANL2012}, we calculate the
consensus amino acid at each position in the HIV genome; (ii) we count how often
each codon coding for the consensus amino acid appears in the MSA; (iii) at each
amino acid position, we divide by the number of sequences in the MSA that had
the consensus amino acid at that position, obtaining {\it codon frequencies}
$\nu_c$; (iv) we calculate the codon entropy from each position as: $S := -
\sum_{c} \nu_c \log \nu_c$, where $c$ runs over codons that code for the
consensus amino acid at this site; (v) we divide by the maximal codon entropy of
that amino acid (e.g. $\log 2$ for twofold degenerate codons). All parts of
{\it env} that are part of a different gene (signaling peptide, second {\it rev}
exon) have been excluded from our main analysis, to avoid contamination about
synonymity. Note that all gap-rich columns of the MSA are stripped from this
figure, hence genes such as {\it env} might appear shorter than they actually
are.
}
\label{fig:syndiv_genome}
\end{center}
\end{figure}
\newpage
% 
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section{Nonsynonymous changes outside of variable regions are deleterious}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=0.8\linewidth]{synmut_conservation_4fold_synnonsyn}
% \caption{Cumulative distribution of synonymous and nonsynonymous diversity in
% {\it gag} in the LANL reference panel (filtered
% sequences only, version 2011)~\cite{LANL2012}. Sites such that the consensus codon
% has three synonymous and six nonsynonymous single mutants were used, and the number of observed mutants
% of a certain type over the number of possible mutants is plotted. Nonsynonymous
% changes are observed less often, {\it ergo} are more conserved, than
% synonymous changes. It can be therefore assumed, as mentioned in the main text,
% that non-escape nonsynonymous changes involve a large fitness cost.}
% \label{fig:synnonsyncons}
% \end{center}
% \end{figure}
% \newpage
